Influence of Connectivity on Dung Beetle Communities

### 2025-05-02 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
p2 and 3 and brushed up 4

# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
 - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
# # - \fancyfoot[R]{ }
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  pdf_document:
    includes:
      # in_header: style/preamble.tex
      template: null
      keep_tex: no
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995). It is because of this dynamic

@@

With experimental designs we are able to isolate factors might be contributing to ecological patterns and processes.

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked (2) Are corridors benefiting any one functional trait over another? 
<!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->

@@

All dung beetles were counted and identified to species as described in

>>>>

# date: "\\vspace{-1em}"
header-includes:
 # - \AtBeginDocument{\let\maketitle\relax} # this removes default title section
 - \usepackage{lineno}
# - \linenumbers
 - \usepackage[default]{sourcesanspro}
 - \usepackage{fancyhdr} # modify header into left, center, right
 - \pagestyle{fancy} # cues the use of fancy headers
 - \fancyfoot{} # cues the use of fancy footer
 - \fancyhead[L]{EEC MS Thesis Draft} # month & year on header left 
 - \fancyhead[R]{p. \thepage} # name and page on header right
# # - \fancyfoot[R]{ }
# - \usepackage{setspace}{singlespace} # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  pdf_document:
    includes:
      # in_header: style/preamble.tex
      template: null
      keep_tex: no
      # number_sections: true
# documentclass: style/nsf2
biblio-style: apalike
csl: ecology.csl
---

@@

Corridors have been shown to be an important mechanism for for minimizing negative consequences of fragmentation[@haddad_corridor_2003]. By improving habitat structure to help facilitate dispersal, wildlife corridors inform movement dynamics of local populations and can shape land uses and occupancy (insert richard forman 1995). Because of this dynamic it becomes necessary to understand responses by species compositions at all taxonomic levels and potential trophic cascades resulting from changes in habitat structure and connectivity.

@@

By measuring changes in biodiversity and species richness within experimental designs we are able to isolate factors might be contributing to ecological patterns and processes. Past studies have measured changes in biodiversity for many different taxa [@collins_fragmentation_2017; @graham_impact_2022; @tewksbury_corridors_2002], yet much work is still needed to build a full scope for how organisms are being effected. Furthermore it is important that we expand our knowledge on how composition changes might impact functional diversities and potential implications for the effectiveness of ecosystem services.

@@

Dung beetles have emerged as a model system with which to test spatial ecology hypotheses [@ros_how_2012; @roslin_dung_2000].They are an incredibly well studied group of insects which are well known for driving a multitude of ecosystem functions (Hasan et al. 2024). The removal, breakdown, and burial of animal feces drive important ecosystem interactions provided by dung beetles enhancing nutrient cycling and soil quality, the reduction of breeding sites for parasites, and a reduction in methane emissions from dung [@slade_disentangling_2016; @WOS:000350444300013]
Here, we aim to gain an understanding of how dung beetles, a group of insects well known for strong dispersal ability in order to compete for ephemeral resources(Hanski and Cambefort 1991), interact with corridors in their landscapes. We sampled dung beetle communities in experimental landscapes developed for the express purposes of comparing connected and isolated patches, as well as the effects of patch to edge ratio and distance to edge. To ask the question of (1) how landscape connectivity impacts dung beetle assemblages dung beetles were collected, identified, and counted with the expectation that biodiversity and abundance would be higher in patches connected by corridors. Additionally we asked are corridors benefiting any one functional trait over another? 
<!-- Since our experimental system consists of open habitats amongst a forested matrix, we anticipate that species preferring open areas and generalists may be more common in our sampling. -->

@@

All dung beetles were counted and identified to species as described in @nemes_illustrated_2015-1 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures was dried to equalibrium and weighed for biomass measurements.
